---
title: "YourCast: Software for Simultaneous Time Series Forecasting with Your Assumptions"
author: "Federico Girosi, Gary King, and Konstantin Kashin"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
bibliography: YourCast.bib
vignette: >
  %\VignetteIndexEntry{YourCast}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r setup, echo=FALSE}
# global chunk options
knitr::opts_chunk$set(cache = FALSE, collapse=TRUE, fig.path = "figures/", dev = "pdf")
```

# Introduction
`YourCast` implements the methods for demographic forecasting
discussed in @GirKin08. Please read at least Chapter 1 of the book before attempting
to use `YourCast`.

At its most basic, `YourCast` runs linear regressions, and estimates
the usual quantities of interest, such as forecasts, causal effects,
etc.  The benefit of running `YourCast` over standard linear
regression software comes from the improved performance due to
estimating sets of regressions together in sophisticated ways.

`YourCast` avoids the bias that results from stacking datasets from
separate cross-sections and assuming constant parameters, and the
inefficiency that results from running independent regressions in each
cross-section.  `YourCast` instead allows you to tie the different
regressions together probabilistically in ways consistent with what
you know about the world and your data.  The model allows you to have
different covariates with different meanings measured in different
cross-sections.

For example, one might assume that the separate time series
regressions in neighboring (or "similar") countries are more alike.
Our approach is fully Bayesian, but you need not assume as the
standard Bayesian approach does that the *coefficients* (which
are never observed) in neighboring countries are similar.  `YourCast` 
makes it possible to assume instead that neighboring countries are
similar in their values or trends in the *expected value of the dependent variable*.
This approach is advantageous because prior
knowledge almost always exists about the dependent variable (such as
that the age profile of mortality looks like the Nike swoosh), and the
expected value is always on the same metric even when including
explanatory variables that differ in number or meaning in each
country.

The power of `YourCast` to improve forecasts comes from allowing one
to smooth in many sophisticated ways, in addition to across countries.
You can thus decide whether to smooth over indices that are
geographic, grouped versions of underlying continuous variables (such
as age groups), time, or interactions among these.  For example, you
can assume that, unless contradicted by the data, forecasts should be
relatively smooth over time, or that the forecast time trends should
be similar in adjacent age groups, or even that the differences in
time trends between adjacent age groups stay roughly similar as they
vary over countries.  The model works with time-series-cross-sectional
(TS-CS) data but also data for which the time series varies over more
than one cross-section (TS-CS-CS-CS $\ldots$) data such as log-mortality
over time by age, country, sex, and cause).   The specific notion of
"smoothness" or "similarity" used in `YourCast` is also your
choice.  The assumptions made by the statistical model are therefore
governed your choices, and the sophistication of those assumptions and
the degree to which they match empirical reality are, for the most
part, limited only by what you may know or are willing to assume
rather than arbitrary choices embedded in a mathematical model.  In
our work, we have found that `YourCast` makes it possible to improve
forecasts well beyond that possible with traditional regression (or
autoregression) strategies, although of course we make no promises
about the future except that your performance may vary.

# Installation
`YourCast` requires the current version of `R` as well as the
packages `foreign`, `lattice`, `ggplot2`, `reshape2`, `grid`, and `gridExtra`, which are loaded automatically by `YourCast`.  `YourCast` can be installed from the `CRAN` repository:

```{r install_cran, eval=FALSE}
install.packages("YourCast")
```

You can also install the latest version of the package directly from GitHub using the `devtools` package:
```{r install_git, echo=TRUE, eval=FALSE}
devtools::install_github("IQSS/YourCast")
```

After installation, attach the package as usual:
```{r load, echo=TRUE, eval=TRUE, message = FALSE}
library(YourCast)
```

# User's Guide

YourCast works with multiple data sets in the same model.  Thus, we
require a more complicated data structure than the usual single data
frame used for most statistical models in `R`.  Although you can create
this object yourself and input it directly into the `yourcast()`
function, it is normally easier to use the tools we created to build
this data structure automatically.  To do this requires following
three steps:

1. labeling and organizing the data in a format we
describe
2. running the `yourprep()` function to prepare your
data, create an input object in the format we need; and load it into
`R` 
3. running the `yourcast()` function to generate
forecasts.  

We describe these steps in the following subsections, and
follow it with a detailed example that illustrates all steps from
start to end.

## Data Preparations
`YourCast` operates on time series cross-sectional data indexed by (1)
a time period such as a year, (2) a grouped continuous variable such
as an age group, and (3) a spatial or geographic variable such as
geographic region or country.  To fix ideas, we refer to these as
time, age, and geography, respectively, but obviously they may change
in other applications.  (Either the age or geography indexes, but not
both, may be dropped if desired.)  We require a single dependent
variable, such as mortality rates, to be the same and have the same
meaning for all units [see @GirKin08, Section 8.4]. Covariates may differ in number, meaning, and content across both age and geography.

Thus, `YourCast` analyzes a set of data sets, each defined for one
cross-sectional unit indexed by age and/or geography.  Inside the data
set corresponding to each cross-sectional unit is a time series with
measures on the dependent variable and the covariates observed in this
cross section.  An example would be an annual time series (say
1952-1996) with the dependent variable of mortality rates and several
covariates, all within the cross-section of 15-20 year olds in Uganda.
All cross-sections should have the same time indices (1952-1996),
possibly with some different overlapping observation periods, the same
dependent variable, and covariates that are the same, completely
different, or overlapping from cross-section to cross-section.

All the individual cross-section data sets (each containing a time
series) must be on in a single subdirectory on disk in fixed width
text files (`.txt`), comma separated value files (`.csv`),
or Stata data files (`.dta`).  (Alternatively, they may be in
memory, in your `R` workspace.)  Each file must be named with one
string in three parts: an alphanumeric tag of the user's choice, a
geography code of between zero (for no geography index) and four
digits, and an age group code of between zero (for no age group index)
to four digits.  For example, if you have observations on cancer
deaths for age group 45 (which might represent 45-50 year olds) for
U.S. citizens (e.g., geography code `"2450"`), you may decide to
choose tag `'cancer'`.  We would add a file extension as well and so
if the data are in a plain text file, we put these elements together
and the file name would be `cancer245045.txt`.  

So we can understand your coding scheme, include an extra file in the
same directory called `tag.index.code`, where `tag` is the
actual alphanumeric tag you chose (not the word t-a-g).  For the
example above, the filename would be `cancer.index.code`.  The
contents of the file should be 0--4 letters `g` followed by 0--4
letters `a`.  In the example above, the entire contents of the
file is: `ggggaa`.

Optionally, you may also add files that contain labels for each of the
time, age, and geography codes.  If these are included, they will make
text and graphics output easier to interpret (and they may be useful
documentation for you separate from yourcast).  The files are
`tag.T.names` for time periods, `tag.A.names` for age
groups, and `tag.G.names` for geographic regions, where again
"tag" is your chosen alphanumeric tag.  The contents of each file
should be ASCII text with all valid numerical codes in the first
column and a corresponding label in the second column.  Include column
labels in the first row.  So for geography, the second column might be
country names and the columns would be labeled "region" and
"name".  If the codes are interpretable as is, such as is often the
case for age groups and time periods, then you can omit the
corresponding file.

Finally, if you wish to smooth over geographic regions, which the
`"map"` and `"bayes"` methods allow, you must also include a
file called `tag.proximity.txt` where `tag` is your chosen
alphanumeric tag and `.txt` is used for text files but can
also be used with comma separated files (`.csv`), or Stata data
files (`.dta`).  The larger the proximity score, the more
proximate that pair of countries is in the prior; a zero element means
the two geographic areas are unrelated, and the diagonal is ignored.
Each row of the `proximity` file has three columns, consisting of
geographic codes for two countries followed by a score indicating the
proximity of the two geographic regions; please include column labels.
For convenience, geographic regions that are unrelated (and would have
zero entries in the symmetric matrix) may be omitted from
`proximity`.  In addition, `proximity` may include rows
corresponding to geographic regions not included in the present
analysis.

## Loading in the Data

We load in all the data described in the previous section at once by
using the function `yourprep()`.  The only required argument is
the chosen alphanumeric tag (and the subdirectory name if its not the
working directory).  The program will then attempt to load all files
in that directory beginning with the chosen tag and will ignore the
rest.  Then run the function.  For example:

```{r loaddata, eval=FALSE}
ydata <- yourprep(tag="cancerMales")
```

The output object `ydata` (of class`"yourprep"`) now includes all
the data and associated information needed for making forecasts.

## Making Forecasts

To make forecasts, we require the data object, name the variables with
a standard `R` formula, and the model.  A regression is estimated for
each cross section (tied together by any chosen priors), and so an
explanatory variable listed in the formula is used for a particular
cross-section if it is in the formula and it is present in that
cross-section's data set.  (That is, a variable not measured for a
cross-section is dropped only for that cross-section.)  As an example:

```{r forecast, eval=FALSE}
ylc <- yourcast(formula=log(rspi2/popu2) ~ time, dataobj=dta, model="LC", verbose=FALSE)
```

Finally, the output object `ylc` (of class `"yourcast"`) can be
plotted with function `plot.yourcast(ylc)` summarized with
function `summarize(ylc)`, or accessed directly (use
`names(ylc)` to see the contents).

# Example
We now reconstruct the demo `chp.11.1` from start to finish to
illustrate the capabilities of the `YourCast` software and provide
further illustration to the user on how to take advantage of them.

Most users will not have their data in a format easily readable by
`yourcast()`. Thus for this example we will start with the raw
`.txt` files and take advantage of the `yourprep()` software
designed to help users construct the `'dataobj'` list easily.

We have stored the original files we used to create the `'dataobj'`
returned by typing `data(chp.11.1)` in YourCast's
`'extdata'` folder. You can view all the data files in this folder by typing:

```{r seedata2, eval=TRUE}
dir(system.file("extdata", package = "YourCast"))
```

The function `yourprep()` in the `YourCast` package is designed to
help you turn these raw files into a `'dataobj'` that
`yourcast()` can read. The `yourprep()` function works by
scanning either the working directory or another directory you specify
for files beginning with the tag `'csid'`. In the `'data'`
folder we scanned above, there are several files whose names consist
of the `'csid'` tag and a CSID code in the format we will specify
to the function. These are the labels `yourprep()` needs to be
able to recognize and process the files. All files should have an
extension so that `yourprep()` knows how to read them; currently
the function supports fixed-width `.txt` files, comma-separated
value files, and Stata `.dta` files.

Let's examine the first of these cross section text files,
`csid204500.txt`. As we can see below, this file contains all the
years from the first observed year to the last predicted year, with
missing values replaced by `NA`s. Because it was created in the
`R` software, this file already has the years written in as rownames
in a way that `R` can read (and for this reason has only three column
labels).

```{r csid204500, echo=FALSE, eval=TRUE}
csid204500 <- read.table(system.file("extdata", "csid204500.txt", package = "YourCast"), header=T)
head(csid204500)
```

However, we expect that most users will have input that looks like the
next file in the directory, `csid204505.txt`. Below we can see
that the observation year is an extra variable rather than a rowname.
```{r csid204505, echo=FALSE, eval=FALSE}
csid204505 <- read.table(system.file("extdata", "csid204505.txt", package = "YourCast"), header=T)
head(csid204505)
```

If this is the case, you should set the argument `year.var` to
`TRUE` in `yourprep()`; this will automatically convert the
`'year'` variable to a rowname as long as it is labeled
`'year'`.

The `'data'` directory also includes some of the optional files
that we included in our `'dataobj'` for the `chp.11.1`
demo. The first is `proximity.txt`, a list of proximity scores
for pairs of the geographic units. The second is
`cntry.codes.txt`, a list of all the CSID codes for the
geographic units and their respective labels. We will load these using
arguments in the `yourprep()` function.

We're now ready to run the `yourprep()` function. Since the
function already grabs all files beginning with `'csid'` tag, we
only need to specify the directory where the files are stored and the
names of the optional files, `G.names` and `proximity`. Note
that we have set `year.var=TRUE` since one of our files has the
observation year as a separate variable rather than as the rowname. We set the directory to the `data` subdirectory of the `YourCast` package. Note that if the package is installed in the default `R` library, you can instead set `dpath` to `file.path(.Library,"YourCast","data")`.

```{r yourprep, echo=TRUE, eval=TRUE}
dta <- yourprep(dpath=system.file("extdata", package="YourCast"),
                year.var=TRUE, sample.frame=c(1950,2000,2001,2030),
                G.names="cntry.codes.txt", proximity="proximity.txt",
                verbose=FALSE)
```

We have now created a `'dataobj'` called `dta`. Examining the
`'dataobj'`, we can see that it includes the two required elements,
`'data'` and `'index.code'`, as well as two optional
elements.

```{r examine_yourprep1, echo=TRUE, eval=TRUE}
names(dta)
```

Examining the `'data'` element, we can see that it includes all
the cross section files that were in the `dpath`:

```{r examine_yourprep2, echo=TRUE, eval=TRUE}
names(dta$data)
```

We're now ready for a run of `yourcast()`. The first run of the
program in the `chp.11.1` demo file uses the Lee-Carter
model. This model uses few of the capacities of the `YourCast` package
since it does no smoothing, but is good for a quick run of the
function. Use of the smoothing options can be seen in many of the
demos and is explained the `yourcast()` documentation. The code
below produces an output object called `ylc` that is of class
`'yourcast'`:

```{r yourcast, echo=TRUE, eval=TRUE, message=FALSE}
ylc <- yourcast(formula=log(rspi2/popu2) ~ time, dataobj=dta, model="LC")
```

The main output from the `yourcast()` function is the
`'yhat'` element of the output list, which contains the observed
and predicted values for every cross section. This output is difficult
to appreciate without graphics, but we can get a quick summary of our
run of the function by typing `summary(ylc)`:

```{r summary_yourcast, echo=TRUE, eval=TRUE}
summary(ylc)
```

Here we can see basic information about the output object. More
information not printed automatically is available by typing
`names(summary(ylc))`.

We're now ready to plot the observed and predicted values to study the
model output. This can be done simply by typing `plot(ylc)`, but
we have added a few arguments here to enhance the graphical
output. The argument `title` gives a title for the plots by
describing the dependent variable. The argument
`age.opts` allows us to pass options to the `age` plot in the form of a list object. For example, we can choose to not plot the predicted
`'yhat'` values in-sample through the `insamp.predict=FALSE` option. You can see
more of these options by typing `help(plot.yourcast)`.

```{r plot, eval=TRUE, fig.keep='all', fig.width=10, fig.height=5, out.width='\\textwidth'}
plot(ylc, title="Respiratory Infections", age.opts=list(insamp.predict=FALSE))
```

Since we did not specify which type of plot we wanted, the default combination of
`age` and `time` plots is returned.
However, the plotting function can also do either of these
plots separately, as well as three dimensional age-time plots, total count plots, and life expectancy at birth plots. To see these, we
need to use the `plots` argument. For example, here is a call
for the three dimensional plot:

```{r plot3d, eval=TRUE, fig.width=5, fig.height=5, out.width='\\textwidth'}
plot(ylc,title="Respiratory Infections", plots="threedim")
```

Finally, if your analysis includes a large number of geographic areas
such that viewing output sequentially on the device is inconvenient,
there an option in the plotting function to save the output for each
geographic code as a `.pdf` file in the working directory rather
than printing it to the device window. Just set `print="pdf"` and pass
filename and output directory options to the function as a list object using the `print.args` argument.

This ends the example section of the users guide. Please visit the
help files for individual functions or send an email to the YourCast
listserv if you have problems.

## More Information
We have included demos that provide step-by-step instructions on how
to reproduce to graphs in Chapters 2 and 11 of *Demographic Forecasting*. 
A list of these demos can be found by typing:
  
```{r demo, eval=FALSE}
demo(package="YourCast")
```

You can also access the preassembled
`'dataobj'`s used in these demos directly by typing:

```{r seedata3, eval=FALSE}
data(package="YourCast")
```

To either run the demos or load these datasets, replace the package
name in the respective argument with the name of the demo or dataset
of interest; i.e., `demo(chp.11.1)`.  The next section goes
through this particular demo in detail.

For more information on the statistical methods implemented in this
software, please refer to *Demographic Forecasting*.

# References